The Community Land Model, version 5.1 \\ One-at-a-time Parameter Perturbation Ensemble

### 2024-08-12 ms/ms.tex
overleaf

Relative to CLM5.0, version 5.1 includes minor bug fixes, parameter adjustments \cite{birch2021}, and the implementation of biomass heat storage \cite{swenson2019}. The PPE experiment required additional code modifications to enable systematic variation of the full suite of model parameters (many of which were previously `hard-coded', i.e., written directly as numerical values in the model source code). We utilized the biogeochemical cycling version of CLM, which includes active carbon and nitrogen cycles. This is as opposed to the satellite phenology mode, which is driven by observed canopy properties and thus has fewer internal feedbacks, and was explored in \citeA{dagon2020}. We used the model in land-only mode driven by observed meteorological forcing from the GSWP3v1 reanalysis product (\url{http://hydro.iis.u-tokyo.ac.jp/GSWP3/}), with the crop model turned off.

@@

To manage computational cost we leveraged the Semi-Analytic spin-up mode (SASU) recently implemented within CLM5 \cite{lu2020,liao2023}. This new module utilizes a matrix representation on CLM's biogeochemistry to significantly reduce spin-up time by one order of magnitude \cite{luo2022,liao2023}, and had been previously leveraged for a PPE with the Organizing Carbon and Hydrology in Dynamic Ecosystems (ORCHIDEE) model \cite{huang2018}. Our spin-up protocol featured 20 years in Accelerated Decomposition spin-up mode (see \citeA{lawrence2019} and \citeA{thornton2005} for details), followed by 80 years of SASU, followed by 40 years of Native Dynamic spin-up mode, each cycling over a ten-year forcing dataset (described below). This protocol was designed to achieve sufficiently equilibrated carbon and nitrogen states, while minimizing computational time. This spin-up methodology did not always reach full equilibration of deep soil carbon (beyond 1 meter depth) in very few points at high latitude. Certain inferences about deep soil carbon from this ensemble may therefore subject to uncertainty due to spin-up concerns.

@@

Clustering was based on a subset of 18 CTSM variables (Table \ref{tab:sg}). The MVSC algorithm delineates clusters based on the mean and variability of each variable at each grid cell computed for six 30-year climatology windows (1865-1894, 1895-1924, ... , 1985-2014). Clusters were delineated to equalize the multi-dimensional variance across the user-specified number of groups, $k$. We tested 15 values of $k$, ranging from 10 to 800. Utilizing the ILAMB2.5 benchmarking software \cite{collier2018}, we calculated skill scores to quantify how well each candidate sparsegrid mirrored the full grid output (Supp Figure \ref{supp:ilamb}). We opted for a 400-cluster sparsegrid, to balance computational cost against model fidelity. Because our emphasis is on vegetated regions, we masked out Antarctica within the clustering algorithm, whereby we do not provide any output below 60$^\circ$S.

>>>>

Relative to CLM5.0, version 5.1 includes minor bug fixes, parameter adjustments \cite{birch2021}, and the implementation of biomass heat storage \cite{swenson2019}. The PPE experiment required additional code modifications to enable systematic variation of the full suite of model parameters (many of which were previously `hard-coded', i.e., written directly as numerical values in the model source code). These modifications were incorporated into the main branch of the CTSM codebase for future use (see Open Research Section). We utilized the biogeochemical cycling version of CLM, which includes active carbon and nitrogen cycles. This is as opposed to the satellite phenology mode, which is driven by observed canopy properties and thus has fewer internal feedbacks, and was explored in \citeA{dagon2020}. We used the model in land-only mode driven by observed meteorological forcing from the GSWP3v1 reanalysis product (\url{http://hydro.iis.u-tokyo.ac.jp/GSWP3/}), with the crop model turned off.

@@

To manage computational cost we leveraged the Semi-Analytic spin-up mode (SASU) recently implemented within CLM5 \cite{lu2020,liao2023}. This new module utilizes a matrix representation on CLM's biogeochemistry to significantly reduce spin-up time by one order of magnitude \cite{luo2022,liao2023}, and had been previously leveraged for a PPE with the Organizing Carbon and Hydrology in Dynamic Ecosystems (ORCHIDEE) model \cite{huang2018}. Our spin-up protocol featured 20 years in Accelerated Decomposition spin-up mode (see \citeA{lawrence2019} and \citeA{thornton2005} for details), followed by 80 years of SASU, followed by 40 years of Native Dynamic spin-up mode, each cycling over a ten-year forcing dataset (described in Section below). This protocol was designed to achieve sufficiently equilibrated carbon and nitrogen states, while minimizing computational time. This spin-up methodology did not always reach full equilibration of deep soil carbon (beyond 1 meter depth) in very few points at high latitude. Certain inferences about deep soil carbon from this ensemble may therefore subject to uncertainty due to spin-up concerns.

@@

Clustering was based on a subset of 18 CTSM variables (Table \ref{tab:sg}). The MVSC algorithm delineates clusters based on the mean and variability of each variable at each grid cell computed for six 30-year climatology windows (1865-1894, 1895-1924, ... , 1985-2014). Clusters were delineated to equalize the multi-dimensional variance across the user-specified number of groups, $k$. We tested 15 values of $k$, ranging from 10 to 800. Utilizing the ILAMB2.5 benchmarking software \cite{collier2018}, we calculated skill scores to quantify how well each candidate sparsegrid mirrored the full grid output (Supp Figure Stk). We opted for a 400-cluster sparsegrid, to balance computational cost against model fidelity. Because our emphasis is on vegetated regions, we masked out Antarctica within the clustering algorithm, whereby we do not provide any output below 60$^\circ$S.
